%!TEX root=../../thesis_ESG.tex
\chapter{Deterministic multi-sourcing}
\label{chap:multi-sourcing:deterministic}

\section{Introduction}


\subsection{Motivations}
\label{sec:multi-sourcing:deterministic:introduction:motivations}


Deciding which plants should have the ability to produce a (new) item is a long-term decision that companies have to answer when expanding their product portfolio.
Note that several plants may be able to produce the same item that is why this problem is called \emph{multi-sourcing}.
Because of their costs and of the time needed to implement them, multi-sourcing decisions cannot be easily changed.
Thus, they have a high impact on competitiveness when deciding production planning (mid-term decision) and scheduling (short-term decision).
One could think that giving to each plant the ability to produce every items is the safest solution to ensure future high service level.
However, multi-sourcing the production of every item leads to unnecessary high assignment costs.
The challenge is to find a trade-off between assignment cost and demand satisfaction.


In multi-sourcing problems, we consider a centralized inventory.
In practice, this can take several forms.
One consists in many plants located in different places but whose productions go to a unique warehouse before being send to the stores.
This is the case for one client in the luxury industry.
Another one consists in several assembly lines in one big plants.
In this cases, it is the lines which are given the ability to produce an item.


Moreover, logistic costs are not considered in the problem.
In practice, in the previous organizations, logistic costs are not a big issue compared to multi-sourcing costs.
In more complex organizations, deciding the best logistic network relies on many more constraints and due to size of the problem, it has to be optimize in its own process.


Argon Consulting meets multi-sourcing problem with clients from different industries.
Thus, Argon Consulting aims at getting a versatile model which can be used in many situations and should be not to too dependent on the industry.
We propose a model which captures the essence of the multi-sourcing problem and which can be easily adapted to particular client cases.
% For example, inventory cost at the end end of a period will not be taken into account but a particular client may want to bound it.
% Thus the model have to be able to add this constraint.
% Solving method must also be able to adapt to model modifications.


\subsection{Problem statement}
\label{sec:multi-sourcing:deterministic:introduction:problem_statement}


We consider a set $\plants$ of plants producing a set $\REF$ of items over $\horizon$ periods.
There is an upper bound $\capacity_{pt}$ on the total period production of plant $p$ at period $t$ (summed over all items).
This upper bound is expressed in time unit since it correspond to available working hours.


Giving a plant $p$ the ability to produce an item $i$ has a cost $\affect_p^i$.
This cost is paid once and for all for the whole horizon.
When a plant $p$ is able to produce item $i$, there is an upper (resp. lower) bound $\ub_{pt}^i\ge 0$ (resp. $\lb_{pt}^i\ge 0$) on the production of item $i$ in plant $p$ at period $t$.
The capacity needed (in time units) to produce one unit of item $i$ in plant $p$ in period $t$ is $\rate_{pt}^i>0$.


The production and the inventory of item $i$ (summed over all plants) must satisfy a demand $\demand_t^i$ at the end of period $t$.
When production of an item $i$ is not used to satisfy the demand, it can be stored and incurs no cost.
For each item $i$, there is an initial inventory $\inventory_0^i\in\RR_+$.


The goal is to satisfy the whole demand at minimum cost.


We call this problem the \emph{deterministic multi-sourcing problem}.


\medskip


% Some parameters or their time dependence may seem unusual.
Let's give some explanations.
% First, the lower bound on the production.
First, when a company gives a plant $p$ the ability to produce an item $i$, some of them imposes that production of item $i$ must not be under some threshold $\lb_{pt}^i$ at period $t$.
Is is often to train the workers and in this particular case, these thresholds are usually constant and positive during the first periods after assignment and equal to zero during the last periods.
Second, the capacity needed to produce one item may depends on time.
It is often the case when people are involved in production since they become more skilled as time goes by.
% Finally, as we are interested in multi-sourcing decisions, we do not consider production nor holding costs.
Finally, since multi-sourcing is a long-term decision and since deciding inventory levels is a mid-term decision, comparing assignment costs to holding costs does not make sense.
Thus, we do not consider holding costs although it is easy for an industrial to extend the model by adding a constraint keeping inventory below some threshold.
% However, extending the model by adding holding or production cost or bounding the inventory between each period is straightforward.


% \medskip


% We also define the \emph{uniform deterministic multi-sourcing problem}. Contrary to \cref{chap:PDP:deterministic}, this model does not rely on applications on real dataset but on technical considerations for proofs. In this problem, we set
% $$
% \rate_{pt}^i = 1
% ,\quad
% \lb_{pt}^i = 0
% ,\quad
% \ub_{pt}^i = \capacity_{pt} = 1
% .
% $$


\subsection{Main results}
\label{sec:multi-sourcing:deterministic:introduction:main_results}


We model the deterministic multi-sourcing problem as a mixed integer program in \cref{sec:multi-sourcing:deterministic:model-formulation} and prove that deterministic multi-sourcing problem is $\NP$-hard in the strong sense and give some polynomial cases in \cref{sec:multi-sourcing:deterministic:NP-completeness}.




\section{Bibliography}



Production multi-sourcing can be viewed as a lot-sizing on several parallel machines.
Although it is typically a long term decision and for this reason, uncertainty is often considerate, a vast literature exists for the deterministic case.
Reviews can be found in \citet{Sethi1990,Koste1999,Stevenson2007} even if they are larger than the only multi-sourcing problem.
They are interest in every types of flexibility whereas we concentrate our research on mix flexibility (ability to produce different combinations of products given certain capacity).


As explained in \citet{Fiorotto2018}, deterministic settings may appear in real life.
For example, in semi-conductor manufacturing, pharmaceutical production, packing for yogurt or textile industry, demand is know but machines must be qualified to produce items.
They develop a deterministic lot-sizing problem on multiple parallel machines aiming at minimizing the sum of setup, production, holding and backlog costs.
One main constraint is the maximal flexibility of the configuration.
Their work is based on ``chains'' and they shows than ``long chains'' configurations give results almost as good as complete flexibility.
A previous work by \citet{Ignizio2009} aiming at minimizing the balance between the utilization of the machines shows similar results.
An extensive study of this chains based on simulation can be found in \citet{Muriel2006}.


In previous papers, flexibility is upper bounded and this bound is an input of the model.
The following models address a complete flexibility but the trade-off is made thanks to the different costs in the objective function.
In \citet{Rezaei2008}, each supplier have a part of imperfect items and a trade-off must be found between the revenue of selling good and imperfect items and the other costs (like purchase cost or holding cost).
In \citet{Woarawichai2017}, model also includes logistic costs in the objective.


Finally, \citet{Snyder2006b} do not directly address a multi-sourcing problem.
However, their work aims at anticipating a disruption in the Supply Chain and controlling bad cases values.
They consider multiple facilities and multiple clients.
Each facility have a known probability of disruption.
They aims at minimizing the weighted sum of assignment and reassignment if one or many disruption occurs.


% \citet{Rezaei2008}
% => multi-item and multi-supplier
% => choose supplier and quantity
% => imperfection in received quantity


% \citet{Woarawichai2017}
% => multi-item and multi-supplier
% => choose supplier and quantity
% => model with logistic cost, etc.

% \citet{Muriel2006}
% We develop analytical models and an optimization-based simulation tool to study
% the impact of increasing flexibility on shortages, production variability, component inventories, and order vari-
% ability 
% => simultion tool at the end


% \citet{Ignizio2009}
% => multi-item, multi-machine
% => minimize the gap between machine utilization


% \citet{Snyder2006b}
% => kind of multi-sourcing
% => minimize the assignment at level $r$ \ie at the $r$th closest facility
% => but if possible, demand will be served by the closest one (in he objective also)


\section{Model formulations}
\label{sec:multi-sourcing:deterministic:model-formulation}


In this section, we introduce a mixed integer program which models the deterministic multi-sourcing problem.
We introduce the following decision variables.
The quantity of item $i$ produced at period $t$ by plant $p$ is denoted by $\quantity_{pt}^i$ and the inventory at the end of the period is denoted by $\inventory_t^i$ .
We also introduce a binary variable $\open_p^i$ which takes the value 1 if plant $p$ is given the ability to produce item $i$.


The deterministic multi-sourcing problem can be written as
\begin{subequations}\label{eq:det-multi-sourcing}
  \begin{align+}
    \min\quad & \rlap{$\ds \sum_{i\in\REF}\sum_{p\in\plants}\affect_p^i \open_p^i$}
    \label{eq:det-multi-sourcing:objective}
    \\
    \st\quad & \ds \inventory_t^i = \inventory_{t-1}^i + \sum_{p\in\plants}\quantity_{pt}^i - \demand_t^i && \forall t\in\range{\horizon},\ \forall i\in\REF,
    \label{eq:det-multi-sourcing:inventory-dynamic}
    \\
    & \ds \sum_{i\in\REF}\rate_{pt}^i\quantity_{pt}^i \leq \capacity_{pt} && \forall t\in\range{\horizon},\ \forall p\in\plants,
    \label{eq:det-multi-sourcing:capacity}
    \\
    & \ds \lb_{pt}^i \open_p^i \le \rate_{pt}^i\quantity_{pt}^i \le \ub_{pt}^i \open_p^i && \forall t\in\range{\horizon},\ \forall p\in\plants, \forall i\in\REF,
    \label{eq:det-multi-sourcing:big-M}
    \\
    & \ds \inventory_{pt}^i,\ \quantity_{pt}^i \ge 0 && \forall t\in\range{\horizon},\ \forall p\in\plants, \forall i\in\REF.
    \label{eq:det-multi-sourcing:positivity}
    \\
    & \ds \open_p^i \in \crbracket{0,1} && \forall p\in\plants, \forall i\in\REF,
    \label{eq:det-multi-sourcing:boolean}
  \end{align+}
\end{subequations}


Objective~\eqref{eq:det-multi-sourcing:objective} minimizes the assignment costs.
Constraint~\eqref{eq:det-multi-sourcing:inventory-dynamic} is the inventory balance.
Capacity of each plant is ensured by constraint~\eqref{eq:det-multi-sourcing:capacity}.
Constraint~\eqref{eq:det-multi-sourcing:big-M} is both a ``big-M'' constraint and a bound on the production of each item in each plant.
Note that without loss of generality, we can suppose that $\ub_{pt}^i \le \capacity_{pt}$ for each period $t$, each plant $p$ and each item $i$.


% \medskip


% The simplifications made for the uniform deterministic multi-sourcing problem lead to the following linear program
% \begin{subequations}\label{eq:det-multi-sourcing}
%   \begin{align+}
%     \min\quad & \rlap{$\ds \sum_{i\in\REF}\sum_{p\in\plants}\affect_p^i \open_p^i$}
%     \label{eq:det-multi-sourcing:objective}
%     \\
%     \st\quad & \ds \inventory_t^i = \inventory_{t-1}^i + \sum_{p\in\plants}\quantity_{pt}^i - \demand_t^i && \forall t\in\range{\horizon},\ \forall i\in\REF,
%     \label{eq:det-multi-sourcing:inventory-dynamic}
%     \\
%     & \ds \sum_{i\in\REF}\quantity_{pt}^i \leq 1 && \forall t\in\range{\horizon},\ \forall p\in\plants,
%     \label{eq:det-multi-sourcing:capacity}
%     \\
%     & \ds \quantity_{pt}^i \le \open_p^i && \forall t\in\range{\horizon},\ \forall p\in\plants, \forall i\in\REF,
%     \label{eq:det-multi-sourcing:big-M}
%     \\
%     & \ds \open_p^i \in \crbracket{0,1} && \forall p\in\plants, \forall i\in\REF,
%     \label{eq:det-multi-sourcing:boolean}
%     \\
%     & \ds \quantity_{pt}^i,\ \inventory_t^i \ge 0 && \forall t\in\range{\horizon},\ \forall p\in\plants, \forall i\in\REF.
%     \label{eq:det-multi-sourcing:positivity}
%   \end{align+}
% \end{subequations}



\section{NP-completeness}
\label{sec:multi-sourcing:deterministic:NP-completeness}


% In this section, we prove that the deterministic multi-sourcing problem is $\NP$-hard in the strong sense and give some polynomial cases.
For any fixed values of the $\affect_p^i$'s, the deterministic multi-sourcing problem is $\NP$-hard.
Moreover, looking at the special cases which remains $\NP$-hard and at the polynomial cases, we see that the difficulty of the multi-sourcing problem is captured either by the lower bounds $\lb_{pt}^i$ on production or by the assignment cost $\affect_p^i$.



\begin{thm}\label{thm:deterministic-multi-sourcing:strong-NP-hard}
  Deciding if there is a solution of the deterministic multi-sourcing problem is $\NP$-complete in the strong sense.
\end{thm}


Reducing the 3-partition problem to deterministic multi-sourcing problem, we show that deciding if the deterministic multi-sourcing problem has a solution is $\NP$-hard in the strong sense.
We remind that the 3-partition problem consists in deciding whether a given multiset $\crbracket{a_1,\ldots,a_{3m}}$ of integers can be partitioned into triples that all have the same sum.
This problem is known to be $\NP$-complete in the strong sense (see~\citet{Garey1979}) even if $\frac{B}{4} < a_i < \frac{B}{2}$ for each $i$ with $B=\frac{1}{m}\sum_{i=1}^{3m}a_i$.


\begin{proof}
Let $\crbracket{a_1,\ldots,a_{3m}}$ be an instance of the 3-partition problem such that $\frac{B}{4} < a_i < \frac{B}{2}$ for each $i$ with $B=\frac{1}{m}\sum_{i=1}^{3m}a_i$.
We reduce polynomially this problem to an instance of the deterministic multi-sourcing problem.
We set
$$
  \horizon=1
  ,\quad
  \plants=\range[1]{m}
  ,\quad
  \REF=\range[1]{3m}
$$
$$
  \rate_{p,1}^i=1
  ,\quad
  \demand_1^i=\frac{a_i}{B}
  ,\quad
  \lb_{p,1}^i=\frac{a_i}{B}
  ,\quad
  \ub_{p,1}^i=\capacity_{p,1}=1.
$$
Thus, if we have a solution for the 3-partition problem, finding a solution of the deterministic multi-sourcing problem is straightforward assigning each triple to a plant.

Conversely, if we have a solution of the deterministic multi-sourcing problem, since $\demand_1^i=\frac{a_i}{B}>0$ for each item $i$, each item is assigned to at least one plant (otherwise, solution is infeasible).
Since $\frac{1}{4}\capacity_{p,1}=\frac{1}{4} < \frac{a_i}{B} = \lb_{p,1}^i$ ensures that there are at most three items per plant, then each plant produces exactly 3 items.
% Suppose that a plant $p$ is given the ability to produce at least four items $i_1,\ldots,i_4$.
% Then, production of the plant $p$ is greater or equal to $\lb_{p,1}^{i_1}+\ldots+\lb_{p,1}^{i_1}=\frac{a_{i_1}}{B}+\ldots+\frac{a_{i_4}}{B}>1=\capacity_{p,1}$.
% Thus, each plant produces at most 3 items.
% Hence, each plant produces exactly 3 items.
Plants having the same capacity and sum of plant capacities being equal to sum of demands, each triple has the same sum.
Thus, we get a solution of the 3-partition problem.

The conclusion follows from the fact that the 3-partition problem is $\NP$-complete in the strong sense even if $\frac{B}{4} < a_i < \frac{B}{2}$ for each $i$.
\end{proof}


% \begin{thm}\label{thm:deterministic-multi-sourcing:strong-NP-hard}
% The deterministic multi-sourcing problem is $\NP$-hard in the strong sense.
% \end{thm}



\cref{thm:deterministic-multi-sourcing:strong-NP-hard} prove the strong $\NP$-hardness but the proof require instances of the deterministic multi-sourcing problem with many plants and uses the lower bound on production which may be equal to zero in some real instances.
One can ask for simpler cases.
\cref{prop:deterministic-multi-sourcing:NP-hard:special-cases} gives some special cases which remain $\NP$-hard and \cref{prop:deterministic-multi-sourcing:polynomial-cases} gives some polynomial cases.


\begin{prop}\label{prop:deterministic-multi-sourcing:NP-hard:special-cases}
  The following special cases of the deterministic multi-sourcing problem are $\NP$-hard:
  \begin{enumerate}
    \item\label{item:deterministic-multi-sourcing:NP-hard:without-production-lower-bound}
    the deterministic multi-sourcing problem without lower bounds on production,
    \item\label{item:deterministic-multi-sourcing:NP-hard:2-plants}
    the deterministic multi-sourcing problem with only one period and only two plants.
  \end{enumerate}
\end{prop}


\begin{proof}[Proof of \cref{prop:deterministic-multi-sourcing:NP-hard:special-cases} (\cref{item:deterministic-multi-sourcing:NP-hard:without-production-lower-bound})]
Let $\crbracket{a_1,\ldots,a_{3m}}$ be an instance of the 3-partition problem such that $\frac{B}{4} < a_i < \frac{B}{2}$ for each $i$ with $B=\frac{1}{m}\sum_{i=1}^{3m}a_i$.
We reduce polynomially this problem to an instance of the deterministic multi-sourcing problem without lower bounds on production.
We set
$$
  \horizon=1
  ,\quad
  \plants=\range[1]{m}
  ,\quad
  \REF=\range[1]{3m},
$$
$$
  \affect_p^i=1
  ,\quad
  \demand_1^i=\frac{a_i}{B}
  ,\quad
  \rate_{p,1}^i=1
  ,\quad
  \lb_{p,1}^i=0
  ,\quad
  \ub_{p,1}^i=\capacity_{p,1}=1.
$$
Thus, if we have a solution for the 3-partition problem, finding a solution with cost $\card{\REF}$ of the deterministic multi-sourcing problem is straightforward assigning each triple to a plant.

Conversely, suppose that we have a solution with cost $\card{\REF}$ of the deterministic multi-sourcing problem.
Since $\demand_1^i=\frac{a_i}{B}>0$ for each item $i$, each item is assigned to at least one plant (otherwise, solution is infeasible).
Assignment costs being equal to 1 and solution cost being equal to the number of item, each item is assigned to at most one plant, hence to exactly one plant.
Since $\frac{1}{4}\capacity_{p,1}=\frac{1}{4} < \frac{a_i}{B} = \demand_1^i$ ensures that there are at most three items per plant, then each plant produces exactly 3 items and we get a collection of $m$ triples.
Plants having the same capacity and sum of plant capacities being equal to sum of demands, each triple has the same sum.
Thus, we get a solution of the 3-partition problem.

The conclusion follows from the fact that the 3-partition problem is $\NP$-complete in the strong sense even if $\frac{B}{4} < a_i < \frac{B}{2}$ for each $i$.
\end{proof}


% \begin{prop}\label{thm:deterministic-multi-sourcing:NP-hard:2-plants}
% The deterministic multi-sourcing problem is $\NP$-hard even if there is only one period and only two plants.
% \end{prop}


Reducing the partition problem to the deterministic multi-sourcing problem, we show that the deterministic multi-sourcing problem with only one period and only two plants remains $\NP$-hard;
We remind that the partition problem is the task of deciding whether a given set of positive integers can be partitioned into two subsets that have the same sum.
This problem is known to be $\NP$-complete (see~\citet{Garey1979}).


\begin{proof}[Proof of \cref{prop:deterministic-multi-sourcing:NP-hard:special-cases} (\cref{item:deterministic-multi-sourcing:NP-hard:2-plants})]
Let $\crbracket{a_1,\ldots,a_m}$ be an instance of the partition problem.
We reduce polynomially this problem to an instance of the deterministic multi-sourcing problem.
We set
$$
  \horizon=1
  ,\quad
  \plants=\crbracket{1,2}
  ,\quad
  \REF=\range[1]{m},
$$
$$
  \affect_p^i=1
  ,\quad
  \demand_1^i=\frac{2a_i}{\sum_{i=1}^{m}a_i}
  ,\quad
  \rate_{pt}^i=1
  ,\quad
  \lb_{pt}^i=0
  ,\quad
  \ub_{pt}^i=\capacity_{pt}=1.
$$
Thus, if we have a solution for the partition problem, finding a solution with cost $\card{\REF}$ of the deterministic multi-sourcing problem is straightforward.

Conversely, if we have a solution with cost $\card{\REF}$ of the deterministic multi-sourcing problem, positivity of the $a_i$ ensures that each item is assigned to at least one plant.
Cost of the solution being equal to $\card{\REF}$, each item is assigned to exactly one plant.
Plants having the same capacity and sum of plant capacities being equal to sum of demands, each subset define by the assignment has the same sum.
Thus, we get a solution of the partition problem.

The conclusion follows from the fact that the partition problem is $\NP$-complete.
\end{proof}


We now gives the polynomial cases.

\begin{prop}\label{prop:deterministic-multi-sourcing:polynomial-cases}
The following special cases of the deterministic multi-sourcing problem are polynomial:
\begin{enumerate}
  \item deterministic multi-sourcing problem with a single plant ($\plants=\crbracket{1}$),
  \item deterministic multi-sourcing problem without lower bound on production and without assignment cost ($\lb_{pt}^i=0$ and $\affect_p^i=0$),
  \item deterministic multi-sourcing problem with infinite capacities ($\ub_{pt}^i=\capacity_{pt}=+\infty$).
\end{enumerate}
\end{prop}


\begin{proof}
\emph{Case 1: deterministic multi-sourcing problem with a single plant.}

For each item $i$, we set
$$
\open_1^i=
\left\{
\begin{array}{l}
1\mbox{ if there exists }t\in\range{\horizon}\mbox{ such that }\demand_t^i>0,\\
0\mbox{ otherwise}.
\end{array}
\right.
$$
Then, we solve the resulting linear program to get an optimal solution of the deterministic multi-sourcing problem.
(It may return that the problem is infeasible.)

\medskip

\emph{Case 2: deterministic multi-sourcing problem without lower bound on production and without assignment cost.}

For each plant $p$ and each item $i$, we set $\open_p^i=1$.
These decisions do not affect the cost in the objective function and do not lead to infeasibility since they relax the constraints on inventory and production variables.
Then, we solve the resulting linear program to get an optimal solution of the deterministic multi-sourcing problem.
(It may return that the problem is infeasible.)

\medskip

\emph{Case 3: deterministic multi-sourcing problem with infinite capacity.}

For each item $i$ we choose a plant $p(i)$ among $\argmin_{p\in\plants}\bracket{\affect_p^i}$.
Then, we set
$$
\open_p^i=
\left\{
\begin{array}{l}
1\mbox{ if }p=p(i),\\
0\mbox{ otherwise}.
\end{array}
\right.
$$
Then, we solve the resulting linear program to get an optimal solution of the deterministic multi-sourcing problem.
(It may return that the problem is infeasible.)
\end{proof}


